---
title: "Week 9 Coding Problems ( R: Data Structures)"
subtitle: "STAT 201 Week 9"
author: "Rishi Ramaiya"

format:
  html:
    toc: true
    embed-resources: true
    number-sections: true  # Enables numbering for sections

execute:
  message: false
  warning: false
---



## Practice Exercise 1

USA’s GDP per capita from 1960 to 2021 is given by the vector G in the code chunk below. The values are arranged in ascending order of the year, i.e., the first value is for 1960, the second value is for 1961, and so on.

```{r}
G = c(3007, 3067, 3244, 3375,3574, 3828, 4146, 4336, 4696, 5032,5234,5609,6094,6726,7226,7801,8592,9453,10565,11674,12575,13976,14434,15544,17121,18237,19071,20039,21417,22857,23889,24342,25419,26387,27695,28691,29968,31459,32854,34515,36330,37134,37998,39490,41725,44123,46302,48050,48570,47195,48651,50066,51784,53291,55124,56763,57867,59915,62805,65095,63028,69288)
```

### 
Use matrix to solve the problem


### 
Find the highest GDP per capita for each five-year period from 1961-1965 to 2015-2020

Hint:

* First, create a matrix where each row represents a five-year period of GDP per capita.
* Then, use the apply() function to find the maximum GDP per capita within each period

```{r}
# Define the matrix with 5-year periods
num_years <- length(G)
start_year <- 1960
years <- start_year:(start_year + num_years - 1)


indices <- which(years >= 1961 & years <= 2020)
G_trimmed <- G[indices] 
years_trimmed <- years[indices]

num_periods <- floor(length(G_trimmed) / 5) 
G_matrix <- matrix(G_trimmed[1:(num_periods * 5)], nrow = num_periods, byrow = TRUE)

rownames(G_matrix) <- paste(years_trimmed[seq(1, length(years_trimmed), by = 5)], 
                            years_trimmed[seq(5, length(years_trimmed), by = 5)], sep = "-")

print(G_matrix)

```
### 
Identify the highest GDP per capita in the United States across all the five-year periods defined in the previous question. Additionally, determine the specific five-year period and the exact year in which this maximum GDP per capita occurred.
```{r}
max_GDP_per_period <- apply(G_matrix, 1, max)

max_GDP <- max(max_GDP_per_period)
max_period <- names(which.max(max_GDP_per_period))

max_GDP_index <- which(G == max_GDP)
max_GDP_year <- years[max_GDP_index]



cat("Highest GDP per capita:", max_GDP, "in the period", max_period, "and year", max_GDP_year, "\n")


```
### 
Identify the five-year periods which the GDP per capita experienced a continuous decline each year, indicating a period of financial crisis or economic degradation
```{r}
# Check for continuous decline in each five-year period
declining_periods <- apply(G_matrix, 1, function(x) all(diff(x) < 0))

# Get the names of periods where GDP declined every year
declining_periods_list <- names(declining_periods[declining_periods])

cat("Five-year periods with continuous decline in GDP per capita:", 
    paste(declining_periods_list, collapse = ", "), "\n")

```

## Practice Exercise 2

```{r}
capital_cities <- read.csv('capital_cities.csv')
```

### 
Use `data.frame` to find the top 10 countries closest to DC

1. Print the names of the countries of the top 10 capital cities closest to the US capital - Washington DC. 

2. Create and print a matrix containing the coordinates of the top 10 capital cities closest to Washington DC.

Note that:

1. The *Country Name* for US is given as *United States* in the data.
2. The 'closeness' of capital cities from the US capital is based on the Euclidean distance of their coordinates to those of the US capital.
3. The data includes two records: *United States* and *US Minor Outlying Islands*, both belonging to the U.S., with a distance of zero. These records must be excluded to accurately identify the closest foreign capital cities.

**Hint:**

1. Get the coordinates of Washington DC from `coordinates_capital_cities`. The row that contains the coordinates of DC will have the same index as `United States` has in the vector `country_names`

2. Create a matrix that has coordinates of Washington DC in each row, and has the same number of rows as the matrix `coordinates_capital_cities`.

3. Subtract `coordinates_capital_cities` from the matrix created in (2). Element-wise subtraction will occur between the matrices.

4. Use the `apply()` function on the matrix obtained above to find the Euclidean distance of Washington DC from the rest of the capital cities. 

5. Using the distances obtained above, find the country that has the closest capital to DC (Euclidean distance > 0).

```{r}
# 1. Get the coordinates of Washington DC
dc_row <- capital_cities[capital_cities$`Country.Name` == "United States", ]
# Check if the United States row is found correctly
print(dc_row)

# Extract coordinates of Washington DC (capital of the United States)
dc_coords <- c(dc_row$`Capital.Latitude`, dc_row$`Capital.Longitude`)

# Check if the coordinates are correctly extracted
print(dc_coords)

# 2. Create a matrix with Washington DC's coordinates in each row
dc_matrix <- matrix(rep(dc_coords, nrow(capital_cities)), ncol = 2, byrow = TRUE)

# 3. Subtract coordinates of each capital from Washington DC's coordinates
coord_diff <- cbind(capital_cities$`Capital.Latitude`, capital_cities$`Capital.Longitude`) - dc_matrix

# Check the coord_diff to make sure it's calculated correctly
print(head(coord_diff))

# 4. Compute the Euclidean distances using apply()
distances <- apply(coord_diff, 1, function(x) sqrt(sum(x^2)))

# Check the distances to ensure they're being computed correctly
print(head(distances))

# 5. Exclude distance to Washington DC (which is 0) and find the closest capital
distances <- distances[distances > 0]  # Exclude distance to Washington DC itself (which is 0)
# Check if any distances remain
if (length(distances) > 0) {
  closest_index <- which.min(distances)
  closest_country <- capital_cities$`Country.Name`[closest_index]
  closest_distance <- distances[closest_index]
  
  # Print the result
  cat("The closest capital to Washington DC is", closest_country, "with a distance of", closest_distance, "\n")
} else {
  cat("No valid distances found.\n")
}

```
## Exercise 3

```{r}
movies <- read.csv("movies.csv", stringsAsFactors = FALSE)
```


#### 
What is the datatype of the object `movies`?



#### 
Count the number movies having a negative profit, i.e., their production budget is higher than their worldwide gross. 

Ignore the movies that:

1. Have missing values of production budget or worldwide gross. 

2. Have a zero worldwide gross *(A zero worldwide gross is probably an incorrect value)*. 

```{r}


movies <- read.csv("movies.csv", stringsAsFactors = FALSE)


negative_profit_movies <- movies[!is.na(movies$production_budget) & 
                                !is.na(movies$worldwide_gross) & 
                                movies$worldwide_gross != 0 & 
                                movies$production_budget > movies$worldwide_gross, ]


num_negative_profit <- nrow(negative_profit_movies)


print(num_negative_profit)
```